### 2019-06-02 paper/bw_book.tex
bw book figure and caption done

\documentclass{article}
\usepackage{graphicx}
\begin{document}
\begin{figure}
	\begin{center}
    \includegraphics[width=\textwidth]{bw_book_fig.pdf}
  \end{center}
  \caption{Shown is the data-driven selection and fitting process used in developing a low-energy effective model of the CuO molecule using a density matrix downfolding procedure. The figures on the left indicate the performance of candidate models on sampled data via an R$^2$ score and the degree to which a prior condition is satisfied through a quadratic hinge loss (QHL) as a control parameter $\lambda$ in the cost function is varied. The figure on the right shows the predicted energy E\textsubscript{eff} for the best model, the minimal model at $\lambda$ = 20, against the target energy E\textsubscript{DMC} for training samples used in the regression. All errorbars are 95\% confidence intervals estimated through bootstrappping.}
\end{figure}
\end{document}